О множествах точек на плоскости с целочисленными расстояниями

### 2019-05-23 pps-linear-bound/pps-linear-lower-bound.tex
Drafts of results to be used

>>>>

We will need the following results and notions from~\cite{our-vmmsh-2018}.

\begin{lemma}
	\cite[Corollary 1]{our-vmmsh-2018}
	\label{lem:max_points_on_line}
	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
	Then there is no straight line which contains more than $(d+3)/2$ points of $M$.
\end{lemma}

\begin{lemma}
	\cite[Lemma 4]{our-vmmsh-2018}
	\label{lem:square_container}
	Let $M\in\mathfrak{M}(2,n)$, $\operatorname{diam} M = d$.
	Then $M$ is situated in a square with side length $d$.
\end{lemma}

\begin{definition}
	\textit{Cross} for points $M_1$ and $M_2$, denoted by $cr(M_1,M_2)$, is union of two straight lines:
	the line which the points $M_1$ and $M_2$ are situated on,
	and the bisector of line segment $M_1 M_2$.
\end{definition}

\begin{lemma}
	\label{lem:intervals_cross}
	If \textbf{open} line segments $M_1 M_2$ and $M_3 M_4$ does not intersect,
	$M_1 \neq M_2$, $M_3 \neq M_4$,
	then the set $cr(M_1,M_2) \cap cr(M_3,M_4)$ is either a straight line or contains not more than 4 points.
\end{lemma}

\begin{lemma}
	\label{lemma:quadr_diag_edges}
	Let $ABCD$ be a convex quadrilateral on the plane.
	Then $\max\{AC,BD\}>\min\{AB,BC,CD,DA\}$,
	i.e. at least one diagonal is greater than at least one side.
\end{lemma}

Определим теперь коэффициенты упаковки точек в квадрат $\varphi_k$.
Пусть
\begin{equation*}
	\Phi_k = \{ P \subset [0;1]^2 : |P|=k\}
	,
\end{equation*}
где $[0;1]^2$~--- замкнутый единичный квадрат на плоскости.
Тогда
\begin{equation*}
	\varphi_k = \max_{P \in \Phi_k} \min_{A,B \in P} |A - B|
	.
\end{equation*}
Иначе говоря, через $\varphi_k$ будем обозначать наибольшее число, такое,
что в единичном квадрате нельзя разместить $k$ точек так,
чтобы расстояние между любыми двумя точками было не менее $\varphi_k$.
Проблема отыскания $\varphi_k$ носит название проблемы упаковки точек в квадрат~\cite{locatelli2002packing,costa2013valid}.

\begin{theorem}%[\cite{costa2013valid}]
	\label{thm:varphi_k_bounds}
	\cite{costa2013valid}
	Для $k\geq 2$ выполнено
	\begin{equation*}
		\sqrt{\frac{2}{k\sqrt{3}}}
		\leq
		\varphi_k
		\leq
		\frac{1}{k-1} +
		\sqrt{
			\frac{1}{(k-1)^2}
			+
			\frac{2}{(k-1)\sqrt{3}}
		}
	\end{equation*}
\end{theorem}
